---
title: "(G)ARCH Parameterization"
output:
  html_document:
    df_print: paged
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/neide/OneDrive/Documents/GitHub/GARCH-param')
```

```{r}
source("polymul.r")
source("arch2psi.r")
source("psi2arch.r")
source("sim.arch.r")
source("lik.arch.r")
source("lik.tarch.r")
source("sim.garch.r")
source("lik.garch.r")
source("lik.tgarch.r")
library(tseries)
library(fGarch)
library(rugarch)
```

# ARCH Case

## Check Bijection

```{r}
set.seed(777)
p <- 10
psi <- rnorm(p+1)
out <- cbind(psi,psi2arch(psi),arch2psi(psi2arch(psi)))
colnames(out) <- c("psi","arch","psi")
label <- c("alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
rownames(out) <- label
print(out)
```

## Gaussian Test Simulation

### Simulate and Plot

```{r}
set.seed(111)
T <- 2000
df <- Inf
z.sim <- sim.arch(psi2arch(psi),T,rnorm(p),df)
plot(ts(z.sim))
```

### Test Likelihood Evaluation

```{r}
lik.arch(psi,z.sim,df)
```

###  Fit an ARCH(p) model

```{r}
psi.init <- rep(0,p+1)
arch.fit <- optim(psi.init,lik.arch,data=z.sim,df=Inf,method="BFGS")
print(arch.fit)
out <- cbind(round(psi2arch(arch.fit$par),digits=6),psi2arch(psi))
colnames(out) <- c("MLE","Truth")
label <- c("alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
rownames(out) <- label
print(out)
```
 
### Compare to "tseries garch" Based on Gaussian Error

```{r}
arch.fit2 <- garch(z.sim, order = c(0, p))
```

#### Summarize Output

```{r}
lik.arch(arch2psi(arch.fit2$coef),z.sim,Inf)
out <- cbind(round(arch.fit2$coef,digits=6),psi2arch(psi))
colnames(out) <- c("MLE","Truth")
label <- c("alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
rownames(out) <- label
print(out)
```

### Compare to "fGarch" Based on Gaussian error

Note that we have to put in a numeric value for the order. In this case $p =$ `r p`.

```{r}
arch.fit3 <- garchFit(~ garch(10,0), z.sim)
```

#### Summarize Output

```{r}
lik.arch(arch2psi(arch.fit3@fit$coef[2:(p+2)]),z.sim,Inf)
out <- cbind(round(arch.fit3@fit$coef,digits=6),c(0,psi2arch(psi)))
colnames(out) <- c("MLE","Truth")
label <- c("mu","alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
rownames(out) <- label
print(out)
```

### Compare to "ruGarch" Based on Gaussian error

```{r}
spec <- ugarchspec(variance.model=list(model="sGARCH", garchOrder=c(10,0)),
                   mean.model = list(armaOrder = c(0,0)))
arch.fit4 <- ugarchfit(data = z.sim, spec = spec)
```

#### Summarize Output

```{r}
lik.arch(arch2psi(coef(arch.fit4)[2:(p+2)]),z.sim,Inf)
out <- cbind(round(coef(arch.fit4),digits=6),c(0,psi2arch(psi)))
colnames(out) <- c("MLE","Truth")
label <- c("mu","alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
rownames(out) <- label
print(out)
```



## Student t Test Simulation

### Simulate and Plot

```{r}
set.seed(555)
T <- 2000
df <- 4
z.sim <- sim.arch(psi2arch(psi),T,rnorm(p),df)
plot(ts(z.sim))
```


### Test Likelihood Evaluation

```{r}
lik.arch(psi,z.sim,df)
```

###  Fit an ARCH(p) model
 
```{r}
psi.init <- rep(0,p+2)
arch.fit <- optim(psi.init,lik.tarch,data=z.sim,method="BFGS")
print(arch.fit)
out <- cbind(c(psi2arch(arch.fit$par[1:(p+1)]),2+exp(arch.fit$par[p+2])),c(psi2arch(psi),df))
colnames(out) <- c("MLE","Truth")
label <- c("alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
label <- c(label,"df")
rownames(out) <- label
print(out)
```


### Compare to "tseries garch" Based on Gaussian Error
 
```{r}
arch.fit2 <- garch(z.sim, order = c(0, p))
```


#### Summarize Output

```{r}
lik.arch(arch2psi(arch.fit2$coef),z.sim,Inf)
out <- cbind(round(arch.fit2$coef,digits=6),psi2arch(psi))
colnames(out) <- c("MLE","Truth")
label <- c("alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
rownames(out) <- label
print(out)
```
 

### Compare to "fGarch" Based on Student t error

```{r}
arch.fit3 <- garchFit(~ garch(10,0), data = z.sim, cond.dist = "std")
``` 

#### Summarize Output

```{r}
lik.arch(arch2psi(coef(arch.fit3)[2:(p+2)]),z.sim,coef(arch.fit3)[p+3])
out <- cbind(round(coef(arch.fit3),digits=6),c(0,psi2arch(psi),df))
colnames(out) <- c("MLE","Truth")
label <- c("mu","alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
label <- c(label,"df")
rownames(out) <- label
print(out)
```

### Compare to "ruGarch" Based on Student t error

```{r}
spec <- ugarchspec(variance.model=list(model="sGARCH", garchOrder=c(10,0)),
                   mean.model = list(armaOrder = c(0,0)),
                   distribution.model = "std")
arch.fit4 <- ugarchfit(data = z.sim, spec = spec)
```

#### Summarize Output

```{r}
lik.arch(arch2psi(coef(arch.fit4)[2:(p+2)]),z.sim,coef(arch.fit4)[p+3])
out <- cbind(round(coef(arch.fit4),digits=6),c(0,psi2arch(psi),df))
colnames(out) <- c("MLE","Truth")
label <- c("mu","alpha0")
for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) }
label <- c(label,"df")
rownames(out) <- label
print(out)
```



# GARCH Case

## Check Bijection

```{r}
set.seed(888)
p <- 1
q <- 5
psi <- rnorm(p+q+1)
out <- cbind(psi,psi2arch(psi),arch2psi(psi2arch(psi)))
colnames(out) <- c("psi","arch","psi")
label <- c("alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
rownames(out) <- label
print(out)
```

## Gaussian Test Simulation

### Simulate and Plot

```{r}
set.seed(222)
T <- 2000
df <- Inf
z.sim <- sim.garch(psi2arch(psi),T,rnorm(p),df)
plot(ts(z.sim))
```


### Test Likelihood Evaluation

```{r}
lik.garch(psi,z.sim,p,df)
``` 
 
   
###  Fit a GARCH(p,q) model

```{r}
psi.init <- rep(0,p+q+1)
garch.fit <- optim(psi.init,fn=lik.garch,data=z.sim,p.order=p,df=Inf,method="BFGS")
print(garch.fit)
out <- cbind(round(psi2arch(garch.fit$par),digits=6),psi2arch(psi))
colnames(out) <- c("MLE","Truth")
label <- c("alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
rownames(out) <- label
print(out)
```
 

### Compare to "tseries garch" Based on Gaussian Error

```{r}
garch.fit2 <- garch(z.sim, order = c(q, p))
```  

#### Summarize Output

```{r}
lik.garch(arch2psi(garch.fit2$coef),z.sim,p,df=Inf)
out <- cbind(round(garch.fit2$coef,digits=6),psi2arch(psi))
colnames(out) <- c("MLE","Truth")
label <- c("alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
rownames(out) <- label
print(out)
```

### Compare to "fGarch" Based on Gaussian error

Note that we have to put in a numeric value for the order. In this case $p =$ `r p` and $q =$ `r q`.

```{r}
garch.fit3 <- garchFit(~ garch(1,5), z.sim)
```

#### Summarize Output

```{r}
lik.garch(arch2psi(coef(garch.fit3)[2:(p+q+2)]),z.sim,p,Inf)
out <- cbind(round(coef(garch.fit3),digits=6),c(0,psi2arch(psi)))
colnames(out) <- c("MLE","Truth")
label <- c("mu","alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
rownames(out) <- label
print(out)
```
### Compare to "ruGarch" Based on Gaussian error

```{r}
spec <- ugarchspec(variance.model=list(model="sGARCH", garchOrder=c(1,5)),
                   mean.model = list(armaOrder = c(0,0)))
garch.fit4 <- ugarchfit(data = z.sim, spec = spec)
```

#### Summarize Output

```{r}
lik.garch(arch2psi(coef(garch.fit4)[2:(p+q+2)]),z.sim,p,Inf)
out <- cbind(round(coef(garch.fit4),digits=6),c(0,psi2arch(psi)))
colnames(out) <- c("MLE","Truth")
label <- c("mu","alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
rownames(out) <- label
print(out)
``` 


## Student t Test Simulation

### Simulate and Plot

```{r}
set.seed(444)
T <- 2000
df <- 5
z.sim <- sim.garch(psi2arch(psi),T,rnorm(p),df)
plot(ts(z.sim))
``` 
 

### Test Likelihood Evaluation

```{r}
lik.garch(psi,z.sim,p,df)
```

###  Fit a GARCH(p,q) model
 
```{r}
psi.init <- rep(0,p+q+2)
garch.fit <- optim(psi.init,fn=lik.tgarch,data=z.sim,p.order=p,method="BFGS")
print(garch.fit)
out <- cbind(c(psi2arch(garch.fit$par[1:(p+q+1)]),2+exp(arch.fit$par[p+q+2])),c(psi2arch(psi),df))
colnames(out) <- c("MLE","Truth")
label <- c("alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
label <- c(label,"df")
rownames(out) <- label
print(out)
```  
  
### Compare to "tseries garch" Based on Gaussian Error
 
```{r}
garch.fit2 <- garch(z.sim, order = c(q, p))
```


#### Summarize Output

```{r}
lik.garch(arch2psi(garch.fit2$coef),z.sim,p,df=Inf)
out <- cbind(round(garch.fit2$coef,digits=6),psi2arch(psi))
colnames(out) <- c("MLE","Truth")
label <- c("alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
rownames(out) <- label
print(out)
```
 

### Compare to "fGarch" Based on Student t error

```{r}
garch.fit3 <- garchFit(~ garch(1,5), data = z.sim, cond.dist = "std")
``` 

#### Summarize Output

```{r}
lik.garch(arch2psi(coef(garch.fit3)[2:(p+q+2)]),z.sim,p,coef(garch.fit3)[p+q+3])
out <- cbind(round(coef(garch.fit3),digits=6),c(0,psi2arch(psi),df))
colnames(out) <- c("MLE","Truth")
label <- c("mu","alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
label <- c(label,"df")
rownames(out) <- label
print(out)
```

### Compare to "ruGarch" Based on Student t error

```{r}
spec <- ugarchspec(variance.model=list(model="sGARCH", garchOrder=c(1,5)),
                   mean.model = list(armaOrder = c(0,0)),
                   distribution.model = "std")
garch.fit4 <- ugarchfit(data = z.sim, spec = spec)
```

#### Summarize Output

```{r}
lik.garch(arch2psi(coef(garch.fit4)[2:(p+q+2)]),z.sim,p,coef(garch.fit4)[p+q+3])
out <- cbind(round(coef(garch.fit4),digits=6),c(0,psi2arch(psi),df))
colnames(out) <- c("MLE","Truth")
label <- c("mu","alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
label <- c(label,"df")
rownames(out) <- label
print(out)
```



   
   
# Data Analysis

We examine the *EuStockMarkets* data set of base R (https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/EuStockMarkets), consisting of daily closing prices Of major European stock indices: Germany DAX (Ibis), Switzerland SMI, France CAC, and UK FTSE. The time period is 1991-1998, with weekends and holidays are omitted. 

## Fit Using "tseries garch"

We focus on the DAX log returns.

```{r}
data(EuStockMarkets)  
dax <- diff(log(EuStockMarkets))[,"DAX"]
dax.garch <- garch(dax)  # Fit a GARCH(1,1) to DAX returns
summary(dax.garch)       # ARCH effects are filtered. However, 
plot(dax.garch)          # conditional normality seems to be violated
```

We can examine the likelihood and parameter estimates based on using "tseries garch".

```{r}
p <- 1
q <- 1
lik.garch(arch2psi(dax.garch$coef),dax,p,df=Inf)
out <- round(dax.garch$coef,digits=6)
out <- as.matrix(out)
colnames(out) <- c("MLE")
label <- c("alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
rownames(out) <- label
print(out)
```


## Fit Using t GARCH Stationary Parameterization

We can try using a t distribution with the stationary parameterization. To help out the optimization, we initialize with the Gaussian GARCH(1,1) fit of "tseries garch".

```{r}
psi.init <- c(arch2psi(dax.garch$coef),0)
garch.fit <- optim(psi.init,fn=lik.tgarch,data=dax,p.order=p,method="BFGS")
print(garch.fit)
out <- round(c(psi2arch(garch.fit$par[1:(p+q+1)]),2+exp(garch.fit$par[p+q+2])),digits=6)
out <- as.matrix(out)
colnames(out) <- c("MLE")
label <- c("alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
label <- c(label,"df")
rownames(out) <- label
print(out)
```
   
## Fit Using t GARCH from "fGarch"
   
```{r}
garch.fit3 <- garchFit(~ garch(1,1), data = dax, cond.dist = "std")
``` 

#### Summarize Output

```{r}
lik.garch(arch2psi(coef(garch.fit3)[2:(p+q+2)]),dax,p,coef(garch.fit3)[p+q+3])
out <- cbind(round(coef(garch.fit3),digits=6),NULL)
colnames(out) <- c("MLE")
label <- c("mu","alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
label <- c(label,"df")
rownames(out) <- label
print(out)
```

### Compare to "ruGarch" Based on Student t error

```{r}
spec <- ugarchspec(variance.model=list(model="sGARCH", garchOrder=c(1,1)),
                   mean.model = list(armaOrder = c(0,0)),
                   distribution.model = "std")
garch.fit4 <- ugarchfit(data = dax, spec = spec)
```

#### Summarize Output

```{r}
lik.garch(arch2psi(coef(garch.fit4)[2:(p+q+2)]),dax,p,coef(garch.fit4)[p+q+3])
out <- cbind(round(coef(garch.fit4),digits=6),NULL)
colnames(out) <- c("MLE")
label <- c("mu","alpha0")
if(p > 0) { for(i in 1:p) { label <- c(label,paste("alpha",i,sep="")) } }
if(q > 0) { for(i in 1:q) { label <- c(label,paste("beta",i,sep="")) } }
label <- c(label,"df")
rownames(out) <- label
print(out)
```
 

